---
title: "index"
author: "Renusree"
date: "11/16/2021"
output: html_document
---


```{r setup, include= FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("analysis.R")
library("knitr")

```

## Introduction and Summary Information:

  The prison population and the demographics of the prison population contribute a lot to the the racism we currently see in American society. There have been profound race related trends in the prison population in the US from the 1990s and the dataset I will be exploring with this assignment provides a closer look at how race and region plays a larger role in prison populations. I chose to analyze the following variables during the analysis: 'year','total_pop','black_pop_15to64', 'male_pop_15to64', 'total_jail_pop', 'black_jail_pop', 'male_adult_jail_pop', 'county_name', 'state', 'aapi_jail_pop', 'aapi_pop_15to64', 'latinx_pop_15to64', 'latinx_jail_pop', and 'commuting_zone. Basically, I chose to look at the total population, male population overall and in jail, and different racial populations overall and in jail. I then also chose to look at commuting zone to analyze if that has any effect on prison population. 
  
  
  While exploring the dataset and doing calculations specifically incarceration trends in Washington state, I found that the percent of males in Washington state jails in 1990 was 87.5% and then in 2018 the number went down to 81.25% which wasn't a big difference over the course of 28 years. Afterwards, I went to analyze how race plays a role in jail populations in Washington state and found that in 1990, Washington state's black population was 1.7% and the black jail population was 17.2% which is a disproportionate representation of the black population in jail compared to overall population. I then did the same analysis in 2018 and found that then in Washington state the black population was 3.28% and the black jail population was 15.6% which is still a disproportionate representation but better than the representation in 1990.

```{r}
plot(linechart)
```
## Why did I include a line chart 

I included this line chart because it demonstrates how the Black population, Latino population, and Asian population in jail fluctuated from 1990 to 2018 in Washington state. It specifically shows how black individuals were disproportionately a large percentage of the jail population in 1990, and even in 2018 they consist of the large majority however it has decreased a little bit since 1990. Another pattern that emerged in this chart is that the Latino percentage in jail fluctuated a lot from 1990 to 2018 consistently but the overall general trend is that Latino percentage in jail did not change that much. Lastly, when analyzing the Asian population in jail, there is a general upward trend which is greatly seen in the graph. I chose a line graph to outline these trends as it focuses on showing how multiple different trends change over the same period of time. With this line chart it is easy to understand how the jail percentage for the different racial groups vary because the year change on the x axis is constant for all of them. 

## Including Plots

You can also embed plots, for example:


```{r, echo=FALSE}
plot(scatterplot)
```
  This scatterplot demonstrates how the percentage of the Latino population compares to the percentage of Latino population in jail in Washington state in 1990. This graph evidently shows that there is a large majority of data points when the Latino population is less than 5% but the Latino population in jail is greater than 10% and even close to 15%. In fact, there are a couple data points that show how the percentage of Latino population is less than 20% but the Latino percentage in jail is around 40% which demonstrates how the Latino population are disproportionately represented in the prison population. I decided that a scatterplot is best to represent this trend as it it shows the correlation between two different variables highlighting both the outliers and places where the most data points are concentrated. 



```{r}
plot(incarceration_map)
```
  I included this map to visualize which counties in Washington have a high vs low prison population in the most recent year (2018). From this visualization model, I noticed that King County has the highest jail population in 2018 and some of the Eastern and Western counties have the lowest jail population. I decided that a map would be best to demonstrate this trend as it clearly displays a pattern of which counties have a higher jail population vs lower jail population with a color scheme that is easy for the human eye to read. 



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
